---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.height = 7,
  fig.width = 7
)
```

<!-- badges: start -->
[![R-CMD-check](https://github.com/hypertidy/tissot/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/hypertidy/tissot/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/tissot)](https://CRAN.R-project.org/package=tissot)
<!-- badges: end -->

# tissot

The [Tissot Indicatrix](https://en.wikipedia.org/wiki/Tissot%27s_indicatrix)
characterizes local distortion in map projections. This package computes and
plots indicatrixes using the [PROJ](https://proj.org/) library directly via the
[PROJ R package](https://hypertidy.github.io/PROJ/).

Derived (with permission) from Bill Huber's
[GIS StackExchange answer](https://gis.stackexchange.com/a/5075/482).

## Installation

```r
# install.packages("pak")
pak::pak("tissot")
```

## Quick start

`tissot()` returns a tibble of distortion properties. The second argument
is the projection `target`; `source` defaults to EPSG:4326:

```{r quick}
library(tissot)
tissot(c(147, -42), "+proj=utm +zone=55 +south")
```

Columns include: `scale_h` (meridional), `scale_k` (parallel), `scale_a` /
`scale_b` (max/min singular values), `scale_area`, `angle_deformation`, and
`convergence`.

## Plotting indicatrixes

`indicatrix()` builds plottable ellipses. The dashed circle is the
undistorted reference; the filled ellipse shows the projection's distortion. 

```{r utm55, fig.alt="Tissot indicatrixes on a 5x4 grid across UTM zone 55S, showing near-circular ellipses confirming low distortion near the central meridian at 147°E, with slight elongation toward the zone edges"}
xy <- expand.grid(x = seq(0, 1e6, length.out = 5), y = seq(4900000, 5700000, length.out = 4))
lonlat <- tissot_unproject(xy, source = "+proj=utm +zone=55 +south")
tis <- tissot(lonlat, "+proj=utm +zone=55 +south")
plot(indicatrix(tis), scale = 3e4)
tissot_map()
```

What does that top left indicatrix look like? 

```{r topleft, fig.alt="Single Tissot indicatrix from the top-left corner of the UTM zone 55S grid, showing a slightly tilted ellipse due to convergence away from the central meridian"}
plot(indicatrix(tis)[1])
```

Far from our UTM zone we are in a lot more trouble.

```{r very-bad-trouble, fig.alt=c("Single Tissot indicatrix for UTM zone 55S evaluated at longitude 100°E, far from the central meridian, showing extreme angular and areal distortion", "Tissot indicatrix for Mercator at latitude -42°, showing strong north-south stretching relative to east-west scale", "Tissot indicatrix for Mercator near the equator, showing a near-circular ellipse confirming low distortion close to latitude 0°")}
## UTM zone 55 is at 147 longitude (55 * 6 - 183)
tis <- tissot(cbind(100, -42), "+proj=utm +zone=55 +south")
plot(indicatrix(tis))

##  In Mercator we have well known problems 
tis <- tissot(cbind(147, -42), "+proj=merc")
plot(indicatrix(tis))

## close to the equator Mercator is ok (in exactly the same way that UTM Zone 55 is ok near 147E longitude)
tis <- tissot(cbind(147, 0), "+proj=merc")
plot(indicatrix(tis))

```

Map projection is arbitrary. 


```{r robinson, fig.alt="Grid of Tissot indicatrixes across the Robinson projection with world coastline overlay, showing rounder ellipses near the equator and increasing distortion toward the poles and edges"}
xy <- expand.grid(seq(-150, 150, by = 30), seq(-60, 60, by = 30))
r <- tissot(xy, "+proj=robin")
ii <- indicatrix(r)
plot(ii, scale = 6e5, add = FALSE, show.axes  = TRUE, show.circle = TRUE)
tissot_map()
```

### Distortion summary

```{r summary}
summary(r)
```



## Colour-coded distortion

Pass `fill.by` to colour ellipses by a distortion metric:

```{r coloured, fig.alt="Tissot indicatrixes for the Robinson projection coloured by areal scale factor, with warm colours toward the poles indicating areas where the projection expands area"}
plot(ii, scale = 6e5, add = FALSE, fill.by = "scale_area")
tissot_map()
```

```{r angle_deformation, fig.alt="Tissot indicatrixes for the Robinson projection coloured by angular deformation in degrees, with warm colours in the corners showing where the projection distorts angles most"}
plot(ii, scale = 6e5, add = FALSE, fill.by = "angle_deformation")
tissot_map()
```

## Projection comparison

```{r mollweide, fig.alt="Grid of Tissot indicatrixes for the Mollweide equal-area projection, showing ellipses of consistent area but varying shape across the map"}
m <- tissot(xy, "+proj=moll")
plot(indicatrix(m), scale = 5e5, add = FALSE)
tissot_map()
```

```{r mercator, fig.alt="Grid of Tissot indicatrixes for the Mercator projection, showing circular ellipses near the equator that grow dramatically larger toward the poles while retaining their circular shape"}
merc_xy <- expand.grid(seq(-150, 150, by = 30), seq(-75, 75, by = 15))
me <- tissot(merc_xy, "+proj=merc")
plot(indicatrix(me), scale = 5e5, add = FALSE)
tissot_map()
```

## Rich single-indicatrix plots

A single indicatrix with axes and reference circle:

```{r single2, fig.alt="Single Tissot indicatrix for a Lambert Conformal Conic with standard parallels at -36 and -38 degrees, evaluated at 147°E 42°S, showing a near-circular ellipse indicating low distortion close to the standard parallels"}
ii2 <- indicatrix(c(147, -42), "+proj=lcc +lat_1=-36 +lat_2=-38 +lat_0=-37 +lon_0=145")
plot(ii2[[1]], scale = 1e4, add = FALSE, show.axes = TRUE, show.circle = TRUE)
```


```{r single3, fig.alt="Single Tissot indicatrix for a Lambert Conformal Conic with standard parallels at -15 and +5 degrees, evaluated at 147°E 42°S, showing significant distortion far from the standard parallels"}
ii3 <- indicatrix(c(147, -42), "+proj=lcc +lat_1=-15 +lat_2=5 +lat_0=0 +lon_0=145")
plot(ii3[[1]], scale = 1e4, add = FALSE, show.axes = TRUE, show.circle = TRUE)
```


## Arbitrary projections

Lambert Conformal Conic

```{r lcc, fig.alt="Tissot indicatrixes for a Lambert Conformal Conic projection over the Southern Ocean, coloured by areal scale factor"}
pxy <- expand.grid(seq(100, 200, by = 25), seq(-75, -45, by = 10))
p <- tissot(pxy, "+proj=lcc +lat_0=-60 +lon_0=147 +lat_1=-70 +lat_2=-55")
plot(indicatrix(p), scale = 3e5, add = FALSE, fill.by = "scale_area")
tissot_map()
```

Universal Transverse Mercator

```{r utm, fig.alt="Tissot indicatrixes for UTM zone 55S across a broad Southern Ocean extent, coloured by areal scale factor, showing distortion growing far from the zone"}
qxy <- expand.grid(seq(100, 200, by = 25), seq(-75, -45, by = 10))
p <- tissot(qxy, "EPSG:32755")
plot(indicatrix(p), scale = 3e5, add = FALSE, fill.by = "scale_area")
tissot_map()
```


## Polar projections

In any projection we should refer to a regular grid of points in its crs, else we get weird situations like this, more obvious on an actual pole: 


```{r polar, fig.alt="Tissot indicatrixes for a polar stereographic projection using a regular lon/lat input grid, showing unevenly spaced and heavily distorted ellipses that illustrate the problem with lon/lat sampling in a polar projection"}
polar_xy <- expand.grid(seq(-180, 150, by = 30), seq(-80, -50, by = 10))
p <- tissot(polar_xy, "+proj=stere +lat_0=-90 +lon_0=147")
plot(indicatrix(p), scale = 2.5e5, add = FALSE, fill.by = "scale_area")
tissot_map()
```

```{r laea, fig.alt="Tissot indicatrixes for a polar Lambert Azimuthal Equal Area projection using a regular lon/lat input grid, showing consistent ellipse areas but irregular spacing near the pole"}
la <- tissot(polar_xy, "+proj=laea +lat_0=-90 +lon_0=147")
plot(indicatrix(la), scale = 2.5e5, add = FALSE)
tissot_map()
```

If we push the centre away from the pole in Azimuthal Equidistant, it's useful to see what happens.

```{r aeqd, fig.alt="Tissot indicatrixes for an Azimuthal Equidistant projection centred at 20°S, applied to a polar-region lon/lat grid, showing the distortion pattern that results from an off-pole projection centre"}
lea <- tissot(polar_xy, "+proj=aeqd +lat_0=-20 +lon_0=147")
plot(indicatrix(lea), scale = 2.5e5, add = FALSE)
tissot_map()
```

## Consider generating input in the crs you are assessing

As with the UTM example above, using `tissot_unproject()` it is usually far better to generate a grid in the CRS being assessed. A grid in lon/lat won't be very meaningful in many projections.

```{r polar-proj, fig.alt="Side-by-side Tissot indicatrix plots comparing polar stereographic and Lambert Azimuthal Equal Area projections, both sampled on a regular grid in projected coordinates, showing much more regular and interpretable distortion patterns than a lon/lat grid"}
op <- par(mfrow = c(1, 2))
ext <- c(-180, 150, -80, -50)
crs <- "+proj=stere +lat_0=-90 +lon_0=147"

projext <- reproj::reproj_extent(ext, crs, source = "EPSG:4326")
polar <- expand.grid(seq(projext[1L], projext[2L], by = 30 * 1e5), seq(projext[3L], projext[4L], by = 10 * 1e5))
polar_xy <- tissot_unproject(polar, "EPSG:4326", source = crs)
p <- tissot(polar_xy, crs, source = "EPSG:4326")
plot(indicatrix(p), scale = 2.5e5, add = FALSE, fill.by = "scale_area")
tissot_map()

ext <- c(-180, 150, -80, -50)
crs <- "+proj=laea +lat_0=-90 +lon_0=147"
projext <- reproj::reproj_extent(ext, crs, source = "EPSG:4326")

polar <- expand.grid(seq(projext[1L], projext[2L], by = 30 * 1e5), seq(projext[3L], projext[4L], by = 10 * 1e5))
polar_xy <- tissot_unproject(polar, "EPSG:4326", source = crs)
p <- tissot(polar_xy, crs, source = "EPSG:4326")
plot(indicatrix(p), scale = 2.5e5, add = FALSE, fill.by = "scale_area")
tissot_map()
par(op)
```



## Why this package?

Most "Tissot indicatrix" plots you'll find online are just geographic
circles drawn on the map. They show what happens to a circle under
the projection, which is useful — but it's not the indicatrix.
The indicatrix is the Jacobian of the projection at a point: it gives
you actual scale factors, angular deformation, and areal distortion.
This package computes those.

Other examples: [mgimond](https://mgimond.github.io/tissot/). 


## Code of Conduct

Please note that the tissot project is released with a
[Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.
