---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/readmefigs/README-",
  fig.height = 7,
  fig.width = 7
)
```

<!-- badges: start -->
[![R-CMD-check](https://github.com/hypertidy/tissot/workflows/R-CMD-check/badge.svg)](https://github.com/hypertidy/tissot/actions)
<!-- badges: end -->

# tissot

Compute and visualize the [Tissot indicatrix](https://en.wikipedia.org/wiki/Tissot%27s_indicatrix) for map projections. The indicatrix characterizes local distortion — how a unit circle on the globe is warped into an ellipse by the projection — revealing the trade-offs every map projection makes between preserving shape, area, and direction.

Based on the calculations shared by [Bill Huber on GIS Stack Exchange](https://gis.stackexchange.com/a/5075/482). Uses [gdalraster](https://CRAN.R-project.org/package=gdalraster) for coordinate transformation, so any projection that GDAL/PROJ can handle works out of the box.

## Installation

```r
# install.packages("remotes")
remotes::install_github("hypertidy/tissot")
```

## Quick start

```{r quickstart}
library(tissot)

## Compute distortion properties at a grid of points
xy <- expand.grid(lon = seq(-150, 150, by = 30),
                  lat = seq(-60, 60, by = 30))

r <- tissot(xy, proj.out = "+proj=robin")
r
```

The result is a tibble with one row per input point. Key columns:

| Column | Meaning |
|--------|---------|
| `scale.h` | Meridional (north–south) scale factor |
| `scale.k` | Parallel (east–west) scale factor |
| `scale.a`, `scale.b` | Semi-major and semi-minor axes of the indicatrix ellipse |
| `scale.area` | Areal distortion (= 1 for equal-area projections) |
| `angle_deformation` | Maximum angular distortion in degrees (= 0 for conformal projections) |
| `convergence` | Angle of the projected meridian from grid north |


## Plotting indicatrices

Build indicatrix ellipses and plot them over a map:

```{r robinson}
ii <- indicatrix(r)
plot(ii, scale = 6e5, add = FALSE)
tissot_map()
```

The `indicatrix()` function accepts the `tissot_tbl` directly — it extracts the projection from its attributes. You can also call it with raw coordinates:

```{r single}
ii_single <- indicatrix(130, -42, proj.out = "+proj=stere +lat_0=-90")
plot(ii_single, scale = 3e5, add = FALSE, show.axes = TRUE, show.circle = TRUE)
```


## Comparing projections

```{r comparison, fig.height = 10, fig.width = 7}
xy <- expand.grid(lon = seq(-150, 150, by = 30),
                  lat = seq(-60, 60, by = 30))
projs <- c(Robinson = "+proj=robin",
           Mollweide = "+proj=moll",
           Mercator = "+proj=merc")

par(mfrow = c(3, 1), mar = c(2, 2, 2, 1))
for (nm in names(projs)) {
  r <- tissot(xy, proj.out = projs[nm])
  ii <- indicatrix(r)
  plot(ii, scale = 6e5, add = FALSE, fill.by = "scale.area")
  tissot_map()
  title(main = nm)
}
```

Mollweide is equal-area (uniform colour), Mercator is conformal (circles, but area grows toward the poles), and Robinson is a compromise.


## Colour by distortion

Use `fill.by` to colour-code ellipses by any distortion metric:

```{r fillby}
r <- tissot(expand.grid(seq(-172, 172, by = 15),
                        seq(-82, 82, by = 15)),
            proj.out = "+proj=robin")
ii <- indicatrix(r)

par(mfrow = c(1, 2), mar = c(2, 2, 2, 1))
plot(ii, scale = 6e5, add = FALSE, fill.by = "scale.area")
tissot_map()
title("Areal distortion")

plot(ii, scale = 6e5, add = FALSE, fill.by = "angle_deformation")
tissot_map()
title("Angular deformation")
```


## Rich single-indicatrix plots

Each indicatrix can display the reference unit circle (dashed white), lambda direction (red = parallel scale), and phi direction (blue = meridional scale):

```{r rich}
r <- tissot(147, -42, proj.out = "+proj=lcc +lat_1=-36 +lat_2=-38 +lat_0=-37 +lon_0=145")
ii <- indicatrix(r)
plot(ii[[1]], scale = 2e5, add = FALSE, show.axes = TRUE, show.circle = TRUE)
title("LCC at Hobart")
```


## Polar projections

```{r polar}
polar_pts <- expand.grid(lon = seq(-165, 165, by = 30),
                         lat = seq(-85, -30, by = 10))
p <- tissot(polar_pts, proj.out = "+proj=stere +lat_0=-90 +lon_0=147")
ii <- indicatrix(p)
plot(ii, scale = 5e5, add = FALSE)
tissot_map()
title("Polar Stereographic")
```


## Distortion summary

```{r summary}
r <- tissot(expand.grid(seq(-150, 150, by = 30), seq(-60, 60, by = 30)),
            proj.out = "+proj=robin")
summary(r)
```


## Code of Conduct

Please note that the tissot project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
